Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2018-06-27 abstract.tex
intagrating some gonuke comments

The goal of this work focuses on developing the necessary understanding of
potential misuse of enrichment cascades to design more effective and efficient
international safeguards approaches.  While it is straightforward to design an
enrichment cascade under ideal conditions as a function of the theoretical feed,
product and tails assays, it is very difficult to find reliable information
about the behavior of a given cascade when the feed assay does not match the
design value. Several methods have been developed to assess the behavior of an
enrichment cascade in such circumstances, those methods evaluate the cut, the
feed to product, feed to tail and the product to tail enrichment ratio,
respectively, alpha, beta and gamma, as a function of the cascade feed assay. As
those four parameters depends on each other, determining two of them allows to
compute the other.  The first approach consists of fixing the cut and alpha
recomputing the corresponding assays at each stages of the cascade. The second
one maintains the ideal condition of the cascade (alpha and beta fixed across
the whole cascade), modifying the cut value at each stage accordingly. Both
approaches have been implemented into the Cyclus fuel cycle
simulator\cite{cyclus, mbmore.2018}. The
third fixes the cut and gamma, using both alpha and beta at each stage as free
parameters. The third method has been investigated in \cite{walker.2017}.

>>>>

The goal of this work focuses on developing the necessary understanding of
potential misuse of enrichment cascades. The latter would allow more effective
international safeguards designs and approaches.  While it is straightforward to
design an enrichment cascade under ideal conditions as a function of the
theoretical feed, product and tails assays, it is very difficult to find
reliable information about the behavior of a given cascade when the feed assay
does not match the design value. Several methods have been developed to assess
the behavior of an enrichment cascade in such circumstances. In addition of the
cut ($\theta$) those methods evaluate the feed to product, feed to tails and the
product to tails enrichment ratio, respectively, $\alpha$, $\beta$ and $\gamma$,
as a function of the cascade feed assay. As those four parameters depend on each
other, determining two of them allows to fully define the other.  The first
approach consists of fixing $\theta$ and $\alpha$ recomputing the corresponding
assays at each stages of the cascade. The second one maintains the ideal
condition of the cascade ($\alpha$ and $\beta$ fixed across the whole cascade),
modifying $\theta$ value at each stage accordingly. Both approaches have been
implemented into the Cyclus fuel cycle simulator\cite{cyclus, mbmore.2018}. The
third fixes $\theta$ and $\gamma$, using both $\alpha$ and $\beta$ at each stage
as free parameters. The third method has been investigated in
\cite{walker.2017}.

### 2018-06-27 dis_clusion.tex
intagrating some gonuke comments

It is really interesting to observe that when the cascade is left completely
untouched (Model A) or when it is slightly retuned to maintain the tail to
product enrichment factor as well as the cut of each centrifuges (Model C),
chaining the cascade we can observe high increase of the enrichment at each
level.  On the contrary, when retuning the cut of each centrifuges to maintain
the ideal state of the cascades (Model B) while chaining them, the \gls{HEU}
production rate is favored over the enrichment gain.

The tail recycling allows for each model a huge gain in productivity, even for
then model B where the number of levels required to reach $90w\%$ of $^{235}$U
in the uranium does not change. Even if no cascades chaining allows to retrieve
the same production rate as a direct enrichment, the model B with tail
reprocessing reach about $80\%$ of an optimum production, which is far from
being negligible, allowing to build up a Significant Quantity of \gls{HEU} in
less than 8 months\ldots

@@

This works will be extended to the near future with additional miss-use method,
allowing for example the reconfiguration of the centrifuges in the cascades.

@@

While mathematically correct, the authors do not guaranty the feasibility
different miss-use tuning methods implemented and are welcoming any insight on
the matter.

>>>>

It is really interesting to observe that when the cascade is left completely
untouched (Model A) or when it is slightly retuned to maintain the tails to
product enrichment factor as well as the cut of each centrifuges (Model C),
chaining the cascade we can observe high increase of the enrichment at each
level.  On the contrary, when retuning the cut of each centrifuges to maintain
the ideal state of the cascades (Model B) while chaining them, the \gls{HEU}
production rate is favored over the enrichment gain.

The tails recycling allows for each model a huge gain in productivity, even for
then model B where the number of levels required to reach $90w\%$ of $^{235}$U
in the uranium does not change. Even if no cascades chaining allows to retrieve
the same production rate as a direct enrichment, the model B with tail
reprocessing reach about $80\%$ of an optimum production, which is far from
being negligible, allowing to build up a Significant Quantity of \gls{HEU} in
less than 8 months\ldots

@@

This works will be extended to the near future with additional misuse method,
allowing for example the reconfiguration of the centrifuges in the cascades.

@@

While mathematically correct, the authors do not guaranty the feasibility
different misuse tuning methods implemented and are welcoming any insight on
the matter.

### 2018-06-27 experiement.tex
intagrating some gonuke comments

\section{The experiment}
This works focuses on comparing the different miss-use models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm have been implemented in
the mbmore package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas enrichment cascades.

@@

As mentioned, all the further calculations will be compared to the most
favorable configuration to produce \gls{HEU}, where all the available
centrifuges are used in a single large cascade designed to directly produce
\gls{HEU} from natural uranium, with a tail assay close to $0.3w\%$. The design
characteristic of the reference cascade are summarized in Table
\ref{tab:cascade_config}.

@@

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tail assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to used them, with few or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

@@

Cascade Design &       & Reference  & Default    \\
\midrule
Targeted  & Feed       & $0.71w\%$  & $0.71w\%$  \\
Assays    & Product    & $90w\%$    & $4.0w\%$   \\
          & Tail       & $0.3w\%$   & $0.3w\%$   \\
\midrule
Effective & Product    & $90.35w\%$ & $4.13w\%$  \\
Assays    & Tail       & $0.28w\%$  & $0.28w\%$  \\ 
\midrule
Stages    & Enriching  & 4          & 4          \\
Number    & Stripping  & 39         & 10         \\
\bottomrule
\end{tabular}
  \label{tab:cascade_config}
\end{table}

\subsection{scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tail of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tail flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tail flow when
    tail recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

\subsection{Level population}
In order to assign the optimum number of cascades to each level, a virtual cut as
been computed as:
\begin{equation}
    \theta^{v}_{j} = \frac{F_{j}-N''_{j}}{N'_{j}-N''_{j}},
\end{equation}
where, $j$ represents a level of cascade and $N_{j}$, $N'_{j}$ and $N''_{j}$
respectively the feed, product and tail assay of the cascades at this level.

A flow equation similar to \eqref{eq:flow} is then solve to get the optimum
number of cascade per level. When the tail are not recycled, the $(1-\theta)$
terms are removed from the flow equation.  The results of the level population
are summarized in Table \ref{tab:level}.

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up, to the first level is attibuted the remaining available
cascades.

@@

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the time required for the material to propagate from the
natural uranium source to the n-th level, regardless of the duration of the
timestep (hour, day, month,...). This is also true for the number of timesteps
required to reach the equilibrium assays value in the tail reprocessing cases.
As timestep can be artificially small or large, one will only be considering
equilibrium values flow rates and assays in the following analysis.

>>>>

\section{The experiment}
This works focuses on comparing the different misuse models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm have been implemented in
the $mbmore$ package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas enrichment cascades.

@@

As mentioned, all the further calculations will be compared to the most
favorable configuration to produce \gls{HEU}, where all the available
centrifuges are used in a single large cascade designed to directly produce
\gls{HEU} from natural uranium, with a tails assay close to $0.3w\%$. The design
characteristic of the reference cascade are summarized in Table
\ref{tab:cascade_config}.

@@

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tails assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to used them, with few or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

@@

Cascade Design &       & Reference   & Default    \\
\midrule
Targeted  & Feed       & $0.71w\%$   & $0.71w\%$  \\
Assays    & Product    & $90w\%$     & $3.5w\%$   \\
          & Tails      & $0.3w\%$   & $0.3w\%$   \\
\midrule
Effective & Product    & $90.35w\%$  & $4.13w\%$  \\
Assays    & Tails      & $0.29w\%$   & $0.29w\%$  \\ 
\midrule
Stages    & Enriching  & 4          & 4          \\
Number    & Stripping  & 39         & 10         \\
\bottomrule
\end{tabular}
  \label{tab:cascade_config}
\end{table}

\subsection{scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tails of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

\subsection{Level population}
In order to assign the optimum number of cascades to each level, a virtual cut as
been computed as:
\begin{equation}
    \theta^{v}_{j} = \frac{F_{j}-N''_{j}}{N'_{j}-N''_{j}},
\end{equation}
where, $j$ represents a level of cascade and $N_{j}$, $N'_{j}$ and $N''_{j}$
respectively the feed, product and tails assay of the cascades at this level.

A flow equation similar to \eqref{eq:flow} is then solve to get the optimum
number of cascade per level. When the tails are not recycled, the $(1-\theta)$
terms are removed from the flow equation.  The results of the level population
are summarized in Table \ref{tab:level}.

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up for each level but the first one. To the first level is
attributed the remaining available cascades.

@@

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the time required for the material to propagate from the
natural uranium source to the n-th level, regardless of the duration of the
timestep (hour, day, month,...). This is also true for the number of timesteps
required to reach the equilibrium assays value in the tails reprocessing cases.
As timestep can be artificially small or large, one will only be considering
equilibrium values flow rates and assays in the following analysis.

### 2018-06-27 motivation.tex
intagrating some gonuke comments

Gas centrifuge cascades are usually designed to operate in ideal manner, with no
losses in separative work, the most effective way. To achieve such ideal
configuration, the cascade is designed to be fed with a specific feed assay and
produce the target  enrichment while rejecting tails at a fix assay.

With the current international tensions regarding enrichment capabilities, this
work aims to measure the effectiveness of an enrichment cascade when used outside of
its designed scope and quantify the attractiveness of such way to build up
significant amount of \gls{HEU}.

The present work investigates the performance of an enrichment cycle when chaining
gas enrichment cascades tuned for low enrich uranium production from natural
uranium. As literature on the mater is for obvious reason limited, three
behavior models have been implemented and used to evaluate the response of an
enrich cascade when fed with different assays than the design one. This work
takes also advantage of the Cyclus\cite{cyclus} fuel cycle capabilities to
evaluate the assay blending equilibrium.

>>>>

Gas centrifuge cascades are usually designed to operate in an ideal manner, with
no losses in separative work. To achieve such ideal configuration, the cascade
is designed to be fed with a specific feed assay and produce the target
enrichment while rejecting tails at a fixed assay.

With the current international tensions regarding enrichment capabilities, this
work aims to measure the effectiveness of an enrichment cascade when used outside of
its designed scope and quantify the attractiveness of such way to build up
significant quantity of \gls{HEU}.

The present work investigates the performance of an enrichment cycle when chaining
gas enrichment cascades tuned for low enrich uranium production from natural
uranium. As literature on the mater is for obvious reason limited, three
behavior models have been implemented and used to evaluate the response of an
enrich cascade when fed with different assays than the design one. This work
also takes advantage of the Cyclus\cite{cyclus} fuel cycle capabilities to
evaluate the assay values at equilibrium.
